Periodicity of Geodesic Flows on the Necker Cube Surface

### 2017-10-16 newpaper.tex
101517

@@

\def\ZZn{\ZZ/n\ZZ}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{abstract}
The Necker cube surface is an infinite topological surface obtained by the periodic edge-to-edge tiling of unit cubes in $\RR^3$ along a 2-dimensional subspace realized as a Euclidean cone surface with a discrete, countable set of singularities of cone angles $3\pi$ and $\frac{3\pi}{2}$. Its various symmetries allow for the categorization of periodic and drift-periodic straight line trajectories in terms of their initial trajectory angles. Results are obtained from studying the Teichmuller geodesic flow on a branched four-fold cover of the surface, where the group of deck transformations of the cover preserves the trajectory angle properties, which we use to determine periodicity on what is an infinite translation surface. The initial rational angle is obtained by the ratio of relatively prime integer vectors of the form $(a,b)$. When $a,b$ are both odd the flow is periodic, and whenever $a$ or $b$ is even, the flow is drift-periodic. \\
\end{abstract}

@@

\compav{ignore:}
The Necker cube surface, $\mathbf{S}$, is an infinite topological surface obtained by the periodic tiling of unit cubes by $\mathbb{R}^3$ translational symmetries along a 2-dimensional subspace realized as a Euclidean cone surface with a discrete set of singularities of cone angle $3\pi$ and $\frac{3\pi}{2}$. An isometric ``flattening" of $\mathbf{S}$ is obtained by injective piecewise-linear maps onto the planar subspace. The fundamental domain of the surface admits infinitely unit square \emph{holes} centered on the  $(2\mathbb{Z})^2$ lattice in $\mathbb{C}$. The flat surface, $\mathbf{U}$, has rotational and translation symmetries and a unit tangent bundle, $UT(\mathbf{U})$, that is isomorphic to $\mathbb{Z}/\frac{\pi}{2}\mathbb{Z}$ as a consequence of the edge identifications. A \emph{flat} branched cover, $\tilde{\mathbf{U}}$, of the surface is obtained by a quotient topology defined on four copies of $\mathbf{U}$'s fundamental domain and indexed by elements in $\mathbb{Z}/4\mathbb{Z}$ for the purpose of aligning geodesic flows. It is a \emph{formally étale} cover of $\mathbf{U}$ whose unit tangent bundle is $\mathbb{Z}/2\pi\mathbb{Z}$ and obtained by the monodromy group of the covering map. $\tilde{\mathbf{U}}$ is a translation surface whose quotient by its translational symmetries is a \emph{Veech} surface whose \emph{Veech group} is $SL(2,\mathbb{Z})$-commensurable. This paper's aim is to study the relations between geodesic flows on base surfaces, Deck transformations of their covers, the orbit of rational directions in $\mathbb{Z}^2$ under Veech group action, and use group theoretic results to prove theorems about rational trajectories on the Necker cube surface.
\compav{ignore$\uparrow$}\\
Put your Problem statement here! Example of a Citation\cite[p.219]{Robotics}. Here's Another Citation \cite{Flueck}

The Necker cube[citation] has made numerous appearances throughout history. The artist famous for his use of optical illusions in his work, M.C. Escher, has occasionally used tilings of the Necker cube such as in the lithographs ``Metamorphosis I," and ``Convex and Concave." (Pictured below.) The crystallographer Louis Albert Necker was credited for having extensively studied the Necker cube's geometry[citation] and remarked on this simple, yet fascinating, optical illusion that appears to ``simultaneously protrude from and intrude into the page."[citation:wolfram]\compav{Find new citation} Some might even recognize it as the same surface in which ``Q*bert" hops around on in the 1982 arcade game published by Gottlieb.

\begin{figure}[H]
\begin{center}
\includegraphics[scale=0.4]{escher.jpg}
\includegraphics[scale=0.55]{escher2.jpg}
\caption{The Necker cube tiling as it appears briefly in ``Metamorphosis I," and hung on a banner in ``Convex and Concave."[citation]}
\label{fig:Escher}
\end{center}
\end{figure}

This paper assumes the reader is familiar with fundamental concepts in algebraic topology such as covering space theory, induced homomorphisms on surface homologies of their symmetries, as well as some background in Veech structures of compact translation surfaces. \compav{cite something}
\newpage

\subsection{Background}
We will introduce some objects, definitions, and theorems to be used later in the paper. First and foremost, we need to describe the construction of the Necker cube surface and give it a concrete structure as a subset of $\mathbb{R}^3$.

\noindent Take the following unit squares in $\RR^3$:

\vspace{0.2in}
\begin{tabular}{p{10cm}c}
\begin{align*}
\mathbf{A}_{m,n,p} = [m, m+1]\times[n,n+1]\times\{p\}, 
\\\mathbf{B}_{m,n,p} = \{m+1\}\times[n,n+1]\times[p-1,p],
\\\mathbf{C}_{m,n,p}= [m,m+1]\times\{n+1\}\times[p-1,p].
\end{align*}
&
\shiftleft{0.8in}{\raisebox{-1in}{\includegraphics[scale=1]{label.png}}}
\end{tabular}

\noindent If we take the integer triple $(m,n,p)$ and require that the individual faces are based around them, we obtain the sets

\begin{align*}
\mathbf{A}=\bigcup\big{\{}\mathbf{A}_{m,n,p}: m+n+p=0\big{\}},
\\\mathbf{B}=\bigcup\big{\{}\mathbf{B}_{m,n,p}: m+n+p=0\big{\}},
\\\mathbf{C}=\bigcup\big{\{}\mathbf{C}_{m,n,p}: m+n+p=0\big{\}},
\end{align*}

\noindent and consider them to be piecewise linear maps of infinitely many compact topological spaces, we acquire subsets of the infinite Necker cube surface whose unions cover the surface.

\begin{Def} We denote the \textbf{Necker cube surface} by $\mathbf{S}$, and define it to be the surface of the union of unit squares of the form $\mathbf{S} = \mathbf{A}\cup\mathbf{B}\cup\mathbf{C}$.\end{Def}

The structure can be \emph{flattened} onto plane by piecewise isometric maps onto a subset of $\mathbb C$ with unit squares removed at every even integer pair in the complex plane: 
\begin{figure}[H]
\centering
\includesvg[width=1.1in]{cubesxyzcut} \hspace{0.1in}\raisebox{1.0in}{\text{$\rightarrow$}}\hspace{0.1in}
\raisebox{0.4in}{\includesvg[width=1.5in]{unfoldcut}}
\caption{A piecewise isometry from $\mathbf{S}$ to $\mathbb{C}$. }
\label{fig:Psi}
\end{figure}
\noindent The red/blue dotted lines represent the edges that are split on the plane. The map from one surface to the other is composed of piecewise isometries, $\Psi:\mathbf{S}\rightarrow\RR^3$.

@@

	\left[ \begin{matrix}
	0 & 0 & 1 \\
	0 & 1 & 0 \\
	-1 & 0 & 0
	\end{matrix}\hspace{2mm}\right]
	\left[\begin{array}{c}
		x - \left\lfloor x \right\rfloor
		\\ y- \left\lfloor y \right\rfloor
		\\ -(z - \left\lfloor z \right\rfloor)
		\end{array} \right]
	+
		\left[\begin{array}{c}
			2 \left\lfloor x \right\rfloor - \frac{3}{2}
			\\ 2\left\lfloor y \right\rfloor - \frac{3}{2}
			\\ x - \left\lfloor x \right\rfloor
		\end{array} \right]
			& \text{if } (x,y,z)\in \mathbf{B}	\vspace{2mm}
	\\

		\left[ \begin{matrix}
		1 & 0 & 0 \\
		0 & 0 & 1 \\
		0 & -1 & 0
		\end{matrix}\hspace{2mm}\right]
		\left[\begin{array}{c}
			x - \left\lfloor x \right\rfloor
			\\ y- \left\lfloor y \right\rfloor
			\\ -(z - \left\lfloor z \right\rfloor)
			\end{array} \right]
		+
			\left[\begin{array}{c}
				2 \left\lfloor x \right\rfloor - \frac{3}{2}
				\\ 2\left\lfloor y \right\rfloor - \frac{3}{2}
				\\ y - \left\lfloor y \right\rfloor
			\end{array} \right]
				& \text{if } (x,y,z)\in \mathbf{C}	\vspace{2mm}
\end{cases}
\end{equation}

The flattened surface is contained entirely in $\{(x,y,z)\in\RR~|~z=0\}$, which is isometric to $\CC$. $\mathbf{S}$ is recovered as a topological quotient on the domain

@@

\noindent We denote the relation by $\sim_{\mathbf{P}}$, and the isometric

\begin{Def} $\mathbf U$ is the surface obtained as the topological quotient $\mathbf{P}/\sim_{\mathbf{P}}$, where $\sim_{\mathbf{P}}$ is a minimal relation on $\mathbf{P}$ defined as follows:
\begin{gather*}
\text{Let } x_{0}=(u_{0},v_{0}),x_{1}=(u_{1},v_{1}) \in \mathbf{P}.  \text{ }\sim_{\mathbf{P}} \text { is given as the relation }\\ x_{0}\sim_{\mathbf{P}}x_{1}  \text{ iff } x_{0}=x_{1}
 \\\text{ or, for some }m,n\in\ZZ~x_{0},x_{1} \in {\partial} \left( \left[2m-\frac{1}{2},2m+\frac{1}{2}\right] \times \left[2n-\frac{1}{2},2n+\frac{1}{2}\right] \right)\\
  \left[\begin{array}{c}
u_{1} -2m
\\v_{1}-2n
\end{array}\right] = \left[\begin{matrix}
0 && 1\\
1 && 0
\end{matrix}\right]
\left[ \begin{array}{c}u_{0}-2m\\
v_{0}-2n
\end{array}\right]
. \end{gather*}
\end{Def}

\begin{thm}
The Necker cube surface $\mathbf{S}$ is isometric to $\mathbf{U}$.
\begin{proof}
To show that $\Psi$ is an isometry of the two surfaces, consider $(x,y,z)\in\mathbf{S}$. $\Psi$ is a bijection between sets $\mathbf{P}$ and $\mathbf{S}$ outside of the edges of each cube parallel to the z-axis (Figure $\ref{fig:Psi}$). These edges split into two edges in $\mathbf{P}$, but are identified to be one and the same in $\mathbf{U}$.\\
For $0\leq i\leq 1$, Let $S_1=\{(m+1,n,p-i)\}, S_2=\{(m+1,n+1,p-i)\}, S_3=\{(m,n+1,p-i)\}$ be these edges.
\compav{come back to this.}
\end{proof}
\end{thm}

\begin{figure}[H]
\centering
\includesvg[width=1.8in]{cubesxyzpaths}
\includesvg[width=2.8in]{unfoldcutpath}
\caption{Simple periodic (c) and drift-periodic (a,b) trajectories represented as line segments on $\mathbf{S}$ and $\mathbf{U}$.}
\end{figure}

We see from the figure above that a geodesic flow in what would appear to be the $\frac{\pi}{4}$ direction closes up on the surface. The edges of $\mathbf{U}$ are identified by a $\pm\frac{\pi}{2}$ rotation about a corner of one of these squares. It is a Euclidean cone surface with countably many conic singularities of angle $\frac{3\pi}{2}$ and $3\pi$ (3 singularities per hole). The set of singularities is denoted $Sing(\mathbf{U})$.

\begin{figure}[H]
\centering
\includesvg[width=1.5in.]{quotient}
\caption{A $2\times2$ cut out section centered at each missing square. Edges and verticies identified.}
\label{fig:quotient}
\end{figure}

The surface $\mathbf{U}^0=\mathbf{U}\backslash Sing(\mathbf{U})$ is the surface with these cone singularities \textbf{removed}.

\subsection{Discussion of Results}
This section will detail how an initial trajectory angle is obtained from $\mathbf{S}$, and how it relates to the periodicity of geodesic flow as a statement of the main theorem.

\begin{Def}
Take a trivial straight-line path in $S\subset\mathbb{R}^3$, $\ell:[0,1]\rightarrow\mathbf{S}$ such that for some $m,n,p\in\mathbb{Z},~ \ell(t)$ is contained in exactly one of $\mathbf{A}_{m,n,p},\mathbf{B}_{m,n,p},~or~\mathbf{C}_{m,n,p}$ for all $t\in[0,1]$. The \textbf{initial trajectory angle} of a geodesic on $\mathbf{S}$ is given as $\theta=Arg(\Psi(\ell(1,s))-\Psi(\ell(0,s)))$, where Arg is the principal argument of a vector in $\mathbb{C}$. The \textbf{initial trajectory} is the vector $\dot{v}=(\cos(\theta),\sin(\theta))\in\mathbb{R}^2$.\\
\compav{Not sure if I should proceed with derivative or map}
\begin{figure}[H]
\centering
\includesvg{vectorplane}
\end{figure}
\end{Def}

\noindent We have two subsets of $\mathbb Z^2$ directions that partition the set of possible rational directions $\theta$.
\begin{Def}
Let $v$ be a unit vector of the form $\frac{1}{k}(x,y)\simeq\mathbb{R}/2\pi\mathbb{Z}$ for $x,y\in\mathbb{Z}$ and $k=\sqrt{x^2+y^2}\in\mathbb{R}$.  We call $v$ an \textbf{odd-odd} vector if x and y are relatively prime integers and both odd. We denote the \textbf{set of all odd-odd directions} $(x,y)\in\mathcal{O}$. We say that $v$ is an \textbf{even-odd} vector if x and y are relatively prime integers, and x is even if and only if y is odd. We denote the \textbf{set of all even-odd directions} $(x,y)\in\mathcal{E}$.
\end{Def}

If $v$ is rational, that is $v=k(x,y)$ for integers x,y, then we commonly refer to the pair $(x,y)$ as the \emph{slope} of a given initial trajectory. For example, the following trajectory modeled on the surface is a closed geodesic on $\mathbf{U}$ with initial trajectory (51,31):

\begin{figure}[H]
\centering
\includegraphics[width=4in.]{closed2.png}
\caption{Geodesic flow on $\mathbf{U}$ modeled using sage-flatsurf.}
\end{figure}

\begin{Def}
A \textbf{geodesic flow} on a topological surface, $X$, is a function $\Phi^{v}_t:UT(X)\times\mathbb{R}\rightarrow UT(X)$ such that $\frac{d\Phi}{dt}=v\in UT(X)$. 
\end{Def}

\noindent Experimental evidence indicates that geodesics in the odd-odd directions are periodic, while geodesics in the even-odd directions are drift-periodic. We will prove the theorem below.

\begin{thm}{\textbf{Dynamics of Geodesic Flow on the Necker cube surface.}}\\ Obtain $\theta$ and $v$ as described in Definition 3. Denote the non-singular unit-speed geodesic flow with initial point $s\in\mathbf{U}^0$ in direction $(\theta\mod{\frac{\pi}{2}})\sim\phi\in UT(\mathbf{U}^0)$ by $\Phi^{\phi}_t:UT(\mathbf{U}^0)\times\mathbb{R}\rightarrow UT(\mathbf{U}^0)$ such that $\frac{d\Phi}{dt}=\phi\in UT(X)$  Then the following is true:
\begin{enumerate}[label=(\roman*)]
\item (Periodic) There exists a $t_0 > 0$ such that $\Phi^{\phi}_{t+t_{0}}(s)=\Phi^{\phi}_{t}(s)$ if and only if $v\in\mathcal{O}$.
\item (Drift-Periodic) There exists a $t_0 > 0$ such that $\Phi^{\phi}_{t+t_{0}}(s)= \Phi^{\phi}_{t}\circ\Phi^{\phi}_{t_0}(s)\neq\Phi^{\phi}_t(s)$ if and only if $v\in\mathcal{E}$.
\end{enumerate}
\end{thm}

>>>>

\def\bU{\mathbf{U}} \def\btU{\tilde{\bU}} \def\bUs{\bU^\circ}
\def\btUos{\btU_0^\circ}
\def\bG{\mathbf{G}} \def\bGs{\mathbf{G}^\circ}
\def\hol{\mathbf{hol}}
\def\bM{\mathbf{M}}

@@

\def\ZZn{\ZZ/n\ZZ}
\def\acts{\rotatebox[origin=c]{-90}{$\circlearrowright$}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{abstract}
We study dynamical properties of geodesic flows on a flat, periodically constructed Euclidean cone surface by obtaining an infinite-type translation cover branched over conical singularities, of which admits a flat induced metric and a connection between any two points in the tangent vector bundle of the surface has trivial holonomy. A unit vector flow on the unit tangent bundle of such a surface is well-defined with a canonical vector representation in $\RR^2$. We study this infinite-type surface as a $\ZZ^2$ cover of a compact Veech surface belonging to the stratum of $\cH(2,2,2,2)$ translation surfaces, and use its $SL(2,\RR)-$commensurable Veech group to prove results that relate directions of flows to the periodicity and ergodicity of their lifts in the cover.
\end{abstract}
\begin{figure}[H]
\centering
\includesvg[width=2.2in]{cubesxyzpaths2}
\label{fig:front}
\caption{Periodic and drift-periodic flows on the Necker cube surface}
\end{figure}

@@

Put your Problem statement here! Example of a Citation\cite[p.219]{Robotics}. Here's Another Citation \cite{Flueck}

Historically, the Necker cube[citation] has made numerous appearances in the work of mathematicians, crystallographers, and scientists interested in human visual systems prior to it being popularized in the works of illusionary artist M.C. Escher (pictured below). The crystallographer Louis Albert Necker was credited for having discovered the optical illusion, and studying its geometry[cite]. The solid presentation of the cube, when rendered as a flat surface with its back faces removed, achieves a similar effect when its three visible faces are shaded in a particular way. We are interested in the infinite tiling of this structure (figure $\ref{fig:front}$), which has the appearance of a rhombile tiling of the Euclidean plane, and refer to it as the Necker cube surface.  Video game fans might even recognize it as the same surface that appears in the 1982 arcade game by Gottlieb, ``Q*bert."

\begin{figure}[H]
\begin{center}
\includegraphics[scale=0.4]{escher.jpg}
\includegraphics[scale=0.55]{escher2.jpg}
\caption{The Necker cube tiling as it appears briefly in ``Metamorphosis I," and impressed into a banner in ``Convex and Concave."[citation]}
\label{fig:Escher}
\end{center}
\end{figure}

Flat, periodic surfaces such as the Necker cube surface are of much interest to differential geometers for two reasons in particular. The first is that, locally, neighborhoods are isometric to the plane, making it remarkably easy to describe parallel transports between two points in tangent vector bundles over the surface by linear group actions. The second is that Riemann surfaces constructed out of infinitely many polyhedra can be \emph{flattened} and contained in the plane in such a way that the metric is preserved. The advantage in this case is that one obtains something akin to infinite billiard tables where a geodesic flows are represented by series of straight line segments in the plane. This paper will use such methods to prove dynamical results about Geodesic flows on the Necker cube surface.

\begin{rem}
Familiarity is assumed on the part of the reader with covering space theory, translation (Veech) surfaces, and their associated Veech groups. For general surveys on these topics: [cite], [cite].
\end{rem}

\subsection{Discussion of Results}
Initial experiments suggested there was a correlation between choice of trajectory angle and periodicity. On a surface composed of infinitely many cubes, there is certainly a flat metric endowed on smaller neighborhoods of the surface that are locally isometric to the plane. Via parallel transports of tangent vectors over the edges of the cubes connecting one face to the next, it is possible to describe geodesic behavior on the Necker cube surface as a series of straight line segments that traverse faces and continuously flow in a fixed direction (see figure $\ref{fig:front}$).
\\\\

\begin{wrapfigure}{l}{2.2in.}
\includesvg{vectorplane}
\end{wrapfigure}
We consider classes of directions obtained from unit vectors in $\RR^3$ by projecting a sufficiently small segment of a flow contained onto a parallel plane. These vectors retain all necessary information about their directions and let us consider rational and irrational angles relative to a choice of basis in $\RR^2$. This is what we refer to as the \emph{initial trajectory angle}.
\\\\
Rational component trajectory angles are then further categorized according to the parities of their components.
\begin{Def}
Let $v$ be a unit vector of the form $\frac{1}{k}(x,y)\in\RR^2$ for $x,y\in\mathbb{Z}$ and $k=\sqrt{x^2+y^2}\in\mathbb{R}$.  We call $v$ an \textbf{odd-odd} vector if x,y are relatively prime and both odd. We denote the \textbf{set of all odd-odd directions} by $\mathcal{O}$. We say that $v$ is an \textbf{even-odd} vector if x and y are relatively prime, and x is even if and only if y is odd. We denote the \textbf{set of all even-odd directions} by $\mathcal{E}$.
\end{Def}
We prove the following is true of all rational directions:

\begin{thm*}
Let $\Phi_t$ be the geodesic flow defined on the unit tangent bundle of the Necker cube surface for $t\in\RR$, and initial point $p=(x_0,v_0)$ with $v_0\in\RR$. Then the following is true:
\begin{enumerate}[label=(\roman*)]
\item $\Phi$ is periodic with period $T\in\RR$ if and only if $v\in\cO$.
\item $\Phi$ is drift-periodic with period $T\in\RR$ if and only if $v\in\cE$.
\end{enumerate}
\end{thm*}

The proof for this theorem can be found in XXX and follows from Theorems YYYY and ZZZZZ.
\\
\compav{theorem about most directions being recurrent?}

\subsection{Acknowledgements}
-Pat Hooper\\
- Vincent Delecroix, Ferrán Valdez, pascal hubert\\
- \\

\newpage

\section{Construction}
This section details how we construct the Necker cube surface as an infinite cover of a compact Riemann surface.
\subsection{Geometric Interpretation}
Consider the following surface:
\begin{figure}[H]
\centering
\includesvg[width=2in.]{neckercube}
\end{figure}

We call this surface $\bG$, the \emph{Necker cube}. $\bG$ is obtained by taking three copies of the unit square and identifying the edges in order to resemble the visible portion of a cube. $\bG$ has a total of three vertices. Every vertex has a cone angle of either $3\pi$ or $\frac{3\pi}{2}$ and are \emph{conical points} (or \emph{singularities}) of the surface. We denote the surface without singularities by $\bGs$. The periodic tiling of $\bG$ will take infinitely many copies of $\bG$ and tile them along the edges labeled with solid triangles. This cover of $\bG$ is the Necker cube surface we denote by $\mathbf{S}$. Likewise, the surface tiled by $\bGs$ we denote by $\mathbf{S}^\circ$\\

\noindent Take the following unit squares in $\RR^3$:\\
\vspace{0.2in}
\begin{tabular}{p{10cm}c}
\begin{align*}
\mathbf{A}_{m,n,p} = [m, m+1]\times[n,n+1]\times\{p\}, 
\\\mathbf{B}_{m,n,p} = \{m+1\}\times[n,n+1]\times[p-1,p],
\\\mathbf{C}_{m,n,p}= [m,m+1]\times\{n+1\}\times[p-1,p].
\end{align*}
&
\shiftleft{0.8in}{\raisebox{-1in}{\includegraphics[scale=1]{label.png}}}
\end{tabular}

\noindent If we take the integer triple $(m,n,p)$ and require that the individual faces based around them  has the triple lie on the plane $x+y+z=0$, $\mathbf{S}$ is broken down into the following sets:

\begin{align*}
\mathbf{A}=\bigcup\big{\{}\mathbf{A}_{m,n,p}: m+n+p=0\big{\}},
\\\mathbf{B}=\bigcup\big{\{}\mathbf{B}_{m,n,p}: m+n+p=0\big{\}},
\\\mathbf{C}=\bigcup\big{\{}\mathbf{C}_{m,n,p}: m+n+p=0\big{\}}.
\end{align*}\\
The sets $\mathbf{A},\mathbf{B},\mathbf{C}$ are contained in $\mathbf{S}$. Their union is also the Necker cube surface.

\begin{figure}[H]
\centering
\includesvg[width=2in]{cubesxyzcut}
\caption{The tiling of $\bG$ where the $\mathbf{A}$ faces are copies of copies of the front-facing square of $\bG$.(Edges that were not used in the tiling are labeled).}
\end{figure}

\subsection{Algebraic Interpretation}
We use a series of cutting and gluing operations to obtain an L-shaped surface homeomorphic to the torus branched at three points:\\
\begin{figure}[H]
\centering
\includesvg[width=3in]{LTorus}
\end{figure}
\noindent Which is then cut along these dotted lines and pieced together to resemble the torus with a unit-square removed from the center:\\
\begin{figure}[H]
\centering
\includesvg[width=3in]{BranchTorus}
\caption{$\bGs$ with 4 independent, closed paths labeled identified}
\end{figure}
The labeled paths are independent from eachother, and a quick calculation shows us that its fundamental group $\pi_1(\bGs)\cong \<a,b,c,d\>$, the free group of four generators (denote identity by $\mathbf{1}$). We describe an infinite cover of $\bGs$ such that the associated deck group is isomorphic to $\ZZ^2$
\begin{Def}
We define $\varphi_1:\pi_1(\bGs)\rightarrow \ZZ^2$ as the group homomorphism on generators $a,b,c,d$: \\$c,d\mapsto(0,0)$\\ $a\mapsto(1,0)$\\$b\mapsto(0,1)$\\ $ab,ba\mapsto(1,1)$.
\end{Def}

\begin{Def}
$\bUs$ is the cover of $\bGs$ such that $\pi_1(\bUs)=\ker\varphi$.
\end{Def}

Our cover of $\bGs$ has the fundamental group, $\pi_1(\bUs)=\<\<a^{-1}b^{-1}ab, c, d\>\>$, the conjugate subgroup of elements that map trivially under $\varphi_1$. The covering map $k_{\varphi_1}:\bUs\hookrightarrow\bGs$ takes every point on $\bUs$ to its modular equivalent under these translational symmetries. $\bUs$ is realized as an infinite $\ZZ^2$-tiling of $\bGs$ in $\CC$:
\begin{figure}[H]
\centering
\includesvg[width=2in]{uprime}
\caption{The infinite surface $\bUs$ embedded in $\CC$ with edges identified.}
\end{figure}

\begin{Def}
$\bU$ is the completion of $\bUs$ that includes the branched vertices. 
\end{Def}

\begin{rem}
The surface $\bG$ is homeomorphic to the torus, and its cover $\bU$ is its universal cover as the paths labeled $c,d$ are only non-trivial when $\bG$ is branched at its three conical points.
\end{rem}

\subsection{Isometry From $\mathbf{S}$ to $\bU$}
The Necker cube surface can be \emph{flattened} onto the plane by piecewise isometric maps onto a subspace of $\RR^3$ with (open) unit squares removed at every even integer pair in the plane, what we claim to be $\bU$.
\begin{figure}[H]
\centering
\includesvg[width=1.1in]{cubesxyzcut} \hspace{0.1in}\raisebox{1.0in}{\text{$\rightarrow$}}\hspace{0.1in}
\raisebox{0.4in}{\includesvg[width=1.5in]{unfoldcut}}
\caption{A piecewise isometry from $\mathbf{S}$ to $\bU$. }
\label{fig:Psi}
\end{figure}
\noindent The red/blue dotted lines represent the edges that are split on the plane. The map from one surface to the other is composed of piecewise isometries, $\Psi:\mathbf{S}\rightarrow\RR^3$.

@@

	\left[ \begin{matrix}
	0 & 0 & 1 \\
	0 & 1 & 0 \\
	-1 & 0 & 0
	\end{matrix}\hspace{2mm}\right]
	\left[\begin{array}{c}
		x - \left\lfloor x \right\rfloor
		\\ y- \left\lfloor y \right\rfloor
		\\ -(z - \left\lfloor z \right\rfloor)
		\end{array} \right]
	+
		\left[\begin{array}{c}
			2 \left\lfloor x \right\rfloor - \frac{3}{2}
			\\ 2\left\lfloor y \right\rfloor - \frac{3}{2}
			\\ x - \left\lfloor x \right\rfloor
		\end{array} \right]
			& \text{if } (x,y,z)\in \mathbf{B}\backslash(\mathbf{A}\cup\mathbf{C})	\vspace{2mm}
	\\

		\left[ \begin{matrix}
		1 & 0 & 0 \\
		0 & 0 & 1 \\
		0 & -1 & 0
		\end{matrix}\hspace{2mm}\right]
		\left[\begin{array}{c}
			x - \left\lfloor x \right\rfloor
			\\ y- \left\lfloor y \right\rfloor
			\\ -(z - \left\lfloor z \right\rfloor)
			\end{array} \right]
		+
			\left[\begin{array}{c}
				2 \left\lfloor x \right\rfloor - \frac{3}{2}
				\\ 2\left\lfloor y \right\rfloor - \frac{3}{2}
				\\ y - \left\lfloor y \right\rfloor
			\end{array} \right]
				& \text{if } (x,y,z)\in \mathbf{C}\backslash(\mathbf{A}\cup\mathbf{B})	\vspace{2mm}
\end{cases}
\end{equation}

The flattened surface is contained entirely in $(x,y,0)\in\RR^3$, which is isometric to $\CC$. $\bU$ is recovered as a topological quotient on the domain

@@

\begin{Def} $\mathbf U$ is the surface obtained as the topological quotient $\mathbf{P}/\sim_{\mathbf{P}}$, where $\CC$ is identified with $\RR^2$ in the usual way and $\sim_{\mathbf{P}}$ is a minimal relation on $\mathbf{P}$ defined as follows:
\begin{gather*}
\text{Let } x_{0}=(u_{0},v_{0}),x_{1}=(u_{1},v_{1}) \in \mathbf{P}.  \text{ }\sim_{\mathbf{P}} \text { is given as the relation }\\ x_{0}\sim_{\mathbf{P}}x_{1}  \text{ iff } x_{0}=x_{1}
 \\\text{ or, for some }m,n\in\ZZ~x_{0},x_{1} \in {\partial} \left( \left[2m-\frac{1}{2},2m+\frac{1}{2}\right] \times \left[2n-\frac{1}{2},2n+\frac{1}{2}\right] \right)\\
  \left[\begin{array}{c}
u_{1} -2m
\\v_{1}-2n
\end{array}\right] = \left[\begin{matrix}
0 && 1\\
1 && 0
\end{matrix}\right]
\left[ \begin{array}{c}u_{0}-2m\\
v_{0}-2n
\end{array}\right]
.\end{gather*}\end{Def}

\begin{rem}
Our map $\Psi$ induces an isometry between $\mathbf{S}$ and $\bU$, and its restriction to the branched surfaces induces an isometry between $\mathbf{S}^\circ$ and $\bUs$. This follows from these piecewise Euclidean transformations that preserves our induced flat metric.
\end{rem}
From here on we use $\bUs$ and $\bU$ instead of $\mathbf{S}$ and $\mathbf{S}^\circ$.

\subsection{Four-Fold Cover}
We denote the \emph{unit tangent bundle} on the surface $\bUs$ by $T^1\bUs$, and construct a cover with trivial holonomy. Initial experiments have shown us that geodesics viewed as discontinuous line segments on $\mathbf{P}$ flow in any one of four directions at a time.
\begin{figure}[H]
\centering
\includesvg[width=1.8in]{cubesxyzpaths}
\includesvg[width=2.8in]{unfoldcutpath}
\caption{Simple periodic (c) and drift-periodic (a,b) trajectories represented as line segments on $\mathbf{S}$ and $\mathbf{U}$.\compav{RELABEL THESE}}
\label{fig:holonomypath}
\end{figure}
Observe that a closed loop on $\bUs$ will act on vectors tangent to the surface by a rotation of $\pm\frac{\pi}{2}$ radians (since the surface is flat and embedded in $\CC$, we work with vectors in $\RR^2$).

\begin{figure}[H]
\centering
\includegraphics[width=2.5in]{monodromy.png}
\caption{Effect a nontrivial loop has on arbitrary basis (i,j).}
\label{fig:loop}
\end{figure}

Let $\omega$ be a flat, Riemannian metric connection on $T^1\bUs$. This is allowable since $\bUs$ is obtained as the cover of a flat Riemann surface. We say that $p=(x_0,v_0)\in T^1\bUs$ is a fixed point in the fiber over $x_0$ in the unit tangent bundle of the surface, with $x_0\in\bUs$ and $v_0\in\RR^2$ being unit. Denote the \emph{holonomy group} of the connection with base point p by $Hol_p(\omega)=\{A\in SO(2,\RR)~:~ p \equiv A\cdot p\}$ where $p$ differs from $A\cdot p$ by some rotation of $v_0$, and the \emph{holonomy bundle}  by $H(p)$, the set of points $p'\in T^1\bUs$ where a closed loop on $\gamma:[0,1]\rightarrow\bUs$ lifts to $\gamma^*:[0,1]\rightarrow T^1\bUs$ such that $p=\gamma^*(0)$ and $p'=\gamma^*(1)$.

\begin{Def}
Define $\varphi_2:\pi_1(\bGs)\rightarrow SO(2,\ZZ)$ as the group homomorphism on generators $a,b,c,d$ such that:
\begin{align*}
a,b\mapsto \begin{pmatrix}1 && 0 \\ 0 && 1\end{pmatrix}=I_2,~~~
c\mapsto \begin{pmatrix}0 && -1 \\ 1 && 0\end{pmatrix}=R,~~~
d\mapsto \begin{pmatrix}0 && 1 \\ -1 && 0\end{pmatrix}=R^3
\end{align*}
\end{Def}
The homomorphism restricted to $\pi_1(\bUs)$ factors through $\pi_1(\bGs)$ and we show that:
\begin{lem}
The group $\varphi_2(\pi_1(\bUs,x_0))=Hol_p(\omega)$.
\begin{proof}
Since $c\mapsto R$ and $d\mapsto R^3$, the image of $\varphi_2$ is generated by $R,R^3$ and isomorphic to $SO(2,\ZZ)$. We see from figure $\ref{fig:holonomypath}$ that parallel transports of vectors by non-trivial paths produce clockwise/counter-clockwise rotations equal to that of of the cone angles of the singularities they loop around, all integer multiples of $\frac{\pi}{2}$ that coincide with paths generated by elements $c,d$. Further, if there is an element in $\pi_1(\bUs,x_0)$ conjugated by $[a,b]$, the effect this would have on holonomy would be trivial as the total cone angle sum around each cut out square is an integer multiple of $2\pi$, which agrees with $a$ and $b$'s trivial images under $\phi_2$.
\end{proof}
\end{lem}

Hence, we obtain a natural monodromy representation with the map $m:\pi_1(\bUs,x_0)\rightarrow SO(2,\ZZ)\cong Hol_p(\omega)$, where for $[\gamma]\in\pi_1(\bUs,x_0)$ we have that $\gamma^*(1)=(x_0,m([\gamma])v_0)$. It follows that since $\omega$ is a flat connection, trivial loops have trivial holonomy and $Hol_p(\omega)$ acts on $H(p)$.

\begin{figure}[H]
\centering
\includegraphics[width=4in.]{closed2.png}
\caption{Geodesic flow on $\mathbf{U}$ modeled using sage-flatsurf.}
\label{fig:complicated}
\end{figure}

\begin{Def}
Let $\btUos$ be the cover of $\bUs$ with fundamental group $\pi_1(\btUos)=\ker\varphi_2|_{\pi_1(\bUs)}\trianglelefteq \pi_1(\bUs)$.
\end{Def}
